Extended Data

### 2022-03-14 doc/manuscript/Dow_manuscript_Nature.Rmd
Update Dow_manuscript_Nature.Rmd
refining results & discussion

<!--LENGTH: The typical length of an article with 3-4 modest display items (figures and tables) is 2000-2500 words (summary paragraph plus body text). We're currently at 2641 words and 1 modest + 2 large display items. Articles typically have 30-50 references (currently at 53)-->

@@

Both canopy foliage phenology and stem growth timing shifted earlier as spring temperatures increased (Fig. `r DOY_timing`, `r EDF_pheno_all`). 
We found a consistent effect of temperature ($T_{max}$ or $T_{min}$) throughout the spring, but the strongest effects on stem growth timing were found using $T_{max}$ during a critical temperature window (CTW). 
CTW was identified by measuring the correlation between weekly $T_{max}$ and $DOY_{25}$ for all combinations of consecutive weeks from January 1 to mean $DOY_{25}$ for each xylem architecture-site combination (`r EDF_climwin`).
The CTW was defined as the weeks for which this correlation was strongest, and mean $T_{max}$ during this period (CTW $T_{max}$) used as the independet variable.

@@

For ring- and diffuse- porous species at both sites, warmer CTW $T_{max}$ resulted in earlier achievement of seasonal milestones.
Consistent with findings from previous studies[<!--@parmesan_globally_2003-->;<!--@ibanez_forecasting_2010-->;<!--@friedl_tale_2014-->], leaf phenological milestones advanced at both sites (Fig. `r DOY_timing`a-b, `r EDT_phenology_parameters`), with greenup (DOY when EVI2 first crossed 15% of the segment EVI2 amplitude) advancing 3.45 days/$^\circ$C at SCBI (p=0.016) and 2.4 days/ $^\circ$C at Harvard Forest (p=0.1).<!--Note to Cam: double check that these were updated-->
Similarly, with the exception of ring porous $DOY_{75}$ at SCBI, the stem growth milestones $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$ decreased with CTW $T_{max}$ (Figs. `r schematic_summary`, `r DOY_timing`c-d; `r EDF_pheno_all`).
Specifically, $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$ advanced 0-1.7 days/ $^\circ$C for ring porous species and 2.8-2.9 days/ $^\circ$C for diffuse porous species at SCBI, and 10.3-12.3 days/ $^\circ$C for ring porous species and 0.9-4.2 days/ $^\circ$C for diffuse porous species at Harvard Forest (`r EDT_phenology_parameters`).

@@

In contrast to the pronounced effects of CTW $T_{max}$ on the timing of growth, its effects on $g_{max}$ and $\Delta DBH$ were weak and inconsistent (Figs. `r schematic_summary`, `r EDF_pheno_all`). 
Specifically, $g_{max}$, which occurred very close to $DOY_{50}$, displayed extremely small negative changes (Harvard Forest) or changes in opposite directions (SCBI) in relationship to CTW $T_{max}$ for ring- and diffuse- porous species.
$\Delta DBH$ displayed a weak positive relationship (+0.029-0.08 mm/ $^\circ$C) with CTW $T_{max}$ at SCBI and a weak negative relationship (-0.2-0.03 mm/ $^\circ$C) at Harvard Forest (`r EDF_pheno_Tmax_climwin`).

@@

To understand how annual growth increments have responded to spring temperatures at the centennial scale, we analyzed tree-ring chronologies of `r n_chronologies_SCBI` species at SCBI[@helcoski_growing_2019] and `r n_chronologies_HF` species at Harvard Forest (`r EDT_species`), along with an additional `r n_chronologies-n_chronologies_SCBI-n_chronologies_HF` chronologies from `r n_cores_sites-2` sites (Fig. `r tree_rings`; `r EDF_map`; `r EDT_tree_ring_summary`)[@dorangeville_drought_2018].
In total, our analysis included `r n_chronologies` chronologies representing `r n_cores_species` broadleaf species at `r n_cores_sites` sites distributed from Alabama (34.35$^\circ$N latitude) to Michigan (45.56$^\circ$N latitude) and spanning a 15 $^\circ$C range in April $T_{max}$.
Across all chronologies, the standardized ring-width index (RWI) was significantly (at p $\le$ 0.05) positively correlated with April $T_{max}$ for only 2% of chronologies: 1 of 142 ring porous and 4 of 66 diffuse porous species-site combinations (`r EDT_tree_ring_summary`). 
There were even fewer significant positive correlations with March and May $T_{max}$: 0% and `r round(2/(142+66),2)*100`%, respectively (`r EDT_tree_ring_summary`).
In contrast, RWI was frequently negatively correlated with $T_{max}$ during peak growing season months (May-August), with significant correlations for 52% (May: 45/141, Jun: 107/141, Jul: 91/141, Aug: 53/141) and 46% (May: 10/66, Jun: 52/66, Jul: 36/66, Aug: 23/66) of species-site-month combinations for ring porous and diffuse porous species, respectively.
$T_{min}$ generally exhibited weaker relationships to annual growth than $T_{max}$, with few significant correlations between spring $T_{min}$ and RWI (`r EDF_tree_rings_TMIN`).

To test whether warm springs might result in storage of non-structural carbohydrates (NSC) that would augment growth the following year[@richardson_seasonal_2013], we extended the analysis to examine correlations between RWI and $T_{max}$ in the previous year (`r EDF_tree_rings_TMAX_extended`).
This revealed little correlation between spring temperatures and growth the following year, with significant positive correlations of RWI to March or April $T_{max}$ for 5/142 ring porous chronologies and to April or May $T_{max}$ for 7/66 diffuse porous chronologies.

To test whether the negative effect of summer temperatures might offset an enhancement of growth by warmer spring temperatures, we tested for the joint effects of April and June-July $T_{max}$ on RWI.
Results were qualitatively similar to the univariate correlations (Fig. `r tree_rings`), with significant (at p = 0.05) positive correlations to April $T_{max}$ for only 4% of chronologies and significant negative correlations with June-July $T_{max}$ for 77% of chronologies, supporting that summer temperatures were the more important driver of annual stem growth (`r EDT_tree_ring_summary`).

@@

<!--summary, setting expectation of extra C-->
Together, our results demonstrate that warmer spring temperatures in the temperate deciduous forests of eastern North America advance the timing of stem growth but have little effect on total annual growth (Figs. `r schematic_summary`-`r tree_rings`). 
The observed advance in stem growth under warmer springs parallels phenological advances observed for canopy foliage (Fig. `r DOY_timing`a-b)[@keenan_net_2014;@finzi_carbon_2020;@stuble_plant_2021] and NEE[@keenan_net_2014;@finzi_carbon_2020]. 
However, inconsistent with the concept that an earlier start to growth would increase $ANPP_{woody}$, we demonstrate that warmer springs either hasten the deceleration of stem expansion or otherwise fail to translate extended growing seasons into biologically significant increases in stem growth (Fig. `r schematic_summary`), and thereby have negligible effect on total annual growth for most species and locations (Fig. `r tree_rings`).
Our observations suggest that the deceleration of stem expansion, which occurs mid-summer near the time of peak canopy greenness (`r EDF_growth_curves_all`)[@finzi_carbon_2020;@etzold_number_2022], is driven by cues other than photosynthate limitation, such as water stress, daylength, or sink limitation[@zweifel_why_2021;@etzold_number_2022;@zani_increased_2020;@keenan_timing_2015;<!--@fu_variation_2014;-->@delpierre_wood_2016;@delpierre_temperate_2016].
Combined with observations that warming springs tend to lengthen the season of C sequestration[@lu_no_2022;@finzi_carbon_2020] and increase net C sequestration[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014;@crabbe_extreme_2016;@zhou_explaining_2016;@fu_maximum_2019;@finzi_carbon_2020], our findings imply a lengthening of the period from peak stem growth to the cessation of CO~2~ uptake by the ecosystem and an increase in C allocated to functions other than stem expansion.
What happens with this extra C?

<!--extra C is probably not going to ANPP_woody-->
Additional C fixation in years with warm springs[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014;@crabbe_extreme_2016;@zhou_explaining_2016;@fu_maximum_2019;@finzi_carbon_2020] does not substantially augment stem diameter growth in the current or following year (`r EDF_tree_rings_TMAX_extended`), and is unlikely to augment $ANPP_{woody}$, which, although routinely calculated based on stem growth, can be partially decoupled from it through differences in wood density or C content[@delpierre_wood_2016].
It is theoretically possible that extra C is allocated to enhanced cell wall thickening, a process that lags behind stem expansion[@cuny_woody_2015], or to a higher ratio of high-density latewood to lower-density earlywood, resulting in formation of more C-dense wood in years with warm springs.
However, existing evidence indicates that warm springs have a neutral or negative effect on latewood width[<!--@tardif_influence_2006;-->@roibu_climatic_2020;@kern_multiple_2013], which is more strongly controlled by summer drought stress[<!--@tardif_influence_2006;-->@roibu_climatic_2020;@kern_multiple_2013], suggesting that a positive effect of warm springs on the total C content of annual rings is unlikely. 
Rather, the most likely explanation of the observed lack of correlation between spring temperatures and stem growth is that growth is sink-limited, such that increased GPP does not enhance $ANPP_{woody}$[@delpierre_wood_2016].

@@

<!--.short vs long-term effects-->
It is possible that as spring warming continues, forests will adjust to directional changes in growing season length with an enhancement of $ANPP_{woody}$.
Across latitudinal gradients, warmer spring temperatures are associated with earlier leaf-out[@delgado_differences_2020] and longer growing seasons, which in turn are correlated with greater tree growth[@anderson-teixeira_joint_2021], woody productivity[@banburymorgan_global_2021], and NEE[@churkina_spatial_2005]<!--Can we replace this churkina_2005 ref with another NEE ref? NO, it's making a different point.-->.<!-- The point here is that if you compare across sites in different climates (Harvard Forest and SCBI, for example), you do see greater stem growth associated with the longer growing seasons. Hopefully my rewording makes that clear. -->
Thus, warming spring temperatures are likely to increase the biophysical potential for annual tree growth.
If extra photosynthate made available through a growing difference between GPP and $ANPP_{woody}$ is allocated to functions that relieve limitations on woody growth -- for example, by enhancing nutrient and water acquisition through enhanced allocation to roots[@zohner_how_2021;@liu_phenological_2022] -- it is possible that warming spring temperatures could ultimately increase $ANPP_{woody}$ through indirect mechanisms.
Therefore, understanding how warming spring temperatures are influencing C allocation within ecosystems remains an important outstanding question.

<!--the dominant effect of temperature on ANPP_woody is a negative effect of gs temperature-->
Regardless of the influence of spring temperatures on C cycling within the ecosystem, our results make clear that the dominant effect of temperatures on tree growth occur not in the spring, but during the peak growing season, when increased atmospheric demand associated with high temperatures limits growth[@novick_increasing_2016;@delpierre_wood_2016;@etzold_number_2022;@zweifel_why_2021]. 
Indeed, the timing of peak growth in June and July (`r EDT_phenology_parameters`, `r EDF_growth_curves_all`) coincides with the timing of the greatest sensitivity of annual growth to $T_{max}$ (Fig. `r tree_rings`, `r EDT_tree_ring_summary`).
This finding is consistent with numerous tree-ring studies demonstrating strong sensitivity of growth to drought stress or high temperatures during the peak growing season[@dorangeville_drought_2018;@helcoski_growing_2019;@anderson-teixeira_joint_2021;@babst_twentieth_2019].
<!--*It is also worth noting that growth is often influenced by temperature of the preceding year(s) through its influence on production and storage of NSCs[@refs;@helcoski], and indeed we observe many significant effects of $T_{max}$ in the preceding growing season, albeit very few in the preceding spring (`r EDF_tree_rings_TMAX_extended`).*--><!--this sentence feels tangential, but discussion of these lag effects was requested by one reviewer.-->
Warm springs may also amplify summer drought stress in some times and places, effectively canceling out any positive effects of an extended growing period[@buermann_widespread_2018;@zhang_drought_2021;@lian_summer_2022].
While such an interaction was unlikely to have played a major role within the scope of our dendrometer band study, given relatively mesic conditions and lack of significant correlation between spring temperatures and summer drought stress (see Methods), there is potential for warm springs to have a net negative effect on growth through such mechanisms at other times or places[@gao_earlier_2022].

>>>>

<!--LENGTH: The typical length of an article with 3-4 modest display items (figures and tables) is 2000-2500 words (summary paragraph plus body text). We're currently at 2671 words and 1 modest + 2 large display items. Articles typically have 30-50 references (currently at 53)-->

@@

Both canopy foliage phenology and the timing of stem growth shifted earlier as spring temperatures increased (Fig. `r DOY_timing`, `r EDF_pheno_all`). 
We found a consistent effect of temperature ($T_{max}$ or $T_{min}$) throughout the spring, but the strongest effects on stem growth timing were found using $T_{max}$ during a critical temperature window (CTW). 
CTW was identified by measuring the correlation between weekly $T_{max}$ and $DOY_{25}$ for all combinations of consecutive weeks from January 1 to mean $DOY_{25}$ for each xylem architecture-site combination (`r EDF_climwin`).
The CTW was defined as the weeks for which this correlation was strongest, and mean $T_{max}$ during this period (CTW $T_{max}$) used as our independent variable.

@@

For ring- and diffuse- porous species at both sites, warmer CTW $T_{max}$ resulted in earlier achievement of seasonal milestones.
Consistent with findings from previous studies<!--[@parmesan_globally_2003;--><!--@ibanez_forecasting_2010;--><!--@friedl_tale_2014]-->, leaf phenological milestones advanced at both sites (Fig. `r DOY_timing`a-b, `r EDT_phenology_parameters`), with greenup (DOY when EVI2 first crossed 15% of the segment EVI2 amplitude) advancing 3.45 days/$^\circ$C at SCBI (p=0.016) and 2.4 days/ $^\circ$C at Harvard Forest (p=0.1).<!--Note to Cam: double check that these were updated-->
Similarly, with the exception of ring porous $DOY_{75}$ at SCBI, the stem growth milestones $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$ decreased with CTW $T_{max}$ (Figs. `r schematic_summary`, `r DOY_timing`c-d; `r EDF_pheno_all`).
Specifically, $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$ advanced 0-1.7 days/ $^\circ$C for ring porous species and 2.8-2.9 days/ $^\circ$C for diffuse porous species at SCBI, and 10.3-12.3 days/ $^\circ$C for ring porous species and 0.9-4.2 days/ $^\circ$C for diffuse porous species at Harvard Forest (`r EDT_phenology_parameters`).

@@

In contrast to the pronounced effects of CTW $T_{max}$ on the timing of growth, its effects on $g_{max}$ and $\Delta DBH$ were weak and inconsistent (Fig. `r schematic_summary`, `r EDF_pheno_all`). 
Specifically, $g_{max}$, which occurred very close to $DOY_{50}$, displayed extremely small negative changes (Harvard Forest) or changes in opposite directions (SCBI) in relationship to CTW $T_{max}$ for ring- and diffuse- porous species.
$\Delta DBH$ displayed a weak positive relationship (+0.029-0.08 mm/ $^\circ$C) with CTW $T_{max}$ at SCBI and a weak negative relationship (-0.2-0.03 mm/ $^\circ$C) at Harvard Forest (`r EDF_pheno_Tmax_climwin`).

@@

To understand how annual growth increments have responded to spring temperatures at the centennial scale, we analyzed tree-ring chronologies of `r n_chronologies_SCBI` species at SCBI[@helcoski_growing_2019] and `r n_chronologies_HF` species at Harvard Forest (`r EDT_species`), plus an additional `r n_chronologies-n_chronologies_SCBI-n_chronologies_HF` chronologies from `r n_cores_sites-2` sites (Fig. `r tree_rings`; `r EDF_map`; `r EDT_tree_ring_summary`)[@dorangeville_drought_2018].
In total, our analysis included `r n_chronologies` chronologies representing `r n_cores_species` broadleaf species at `r n_cores_sites` sites distributed from Alabama (34.35$^\circ$N latitude) to Michigan (45.56$^\circ$N latitude) and spanning a 15 $^\circ$C range in April $T_{max}$.
Across all chronologies, the standardized ring-width index (RWI) was significantly (at p $\le$ 0.05) positively correlated with April $T_{max}$ for only 2% of chronologies: 1 of 142 ring porous and 4 of 66 diffuse porous species-site combinations (`r EDT_tree_ring_summary`). 
There were even fewer significant positive correlations with March and May $T_{max}$: 0% and `r round(2/(142+66),2)*100`%, respectively (`r EDT_tree_ring_summary`).
In contrast, RWI was frequently negatively correlated with $T_{max}$ during peak growing season months (May-August), with significant correlations for 52% (May: 45/141, Jun: 107/141, Jul: 91/141, Aug: 53/141) and 46% (May: 10/66, Jun: 52/66, Jul: 36/66, Aug: 23/66) of species-site-month combinations for ring porous and diffuse porous species, respectively.
$T_{min}$ generally exhibited weaker relationships to annual growth than $T_{max}$, with few significant correlations between spring $T_{min}$ and RWI (`r EDF_tree_rings_TMIN`).

To test whether warm springs might result in storage of non-structural carbohydrates (NSC) that would augment growth the following year[@richardson_seasonal_2013], we extended the analysis to examine correlations between RWI and $T_{max}$ in the previous year (`r EDF_tree_rings_TMAX_extended`).
This revealed little effect of previous spring temperatures on annual growth, with significant positive correlations of RWI to previous March or April $T_{max}$ for 5/142 ring porous chronologies and to previous April or May $T_{max}$ for 7/66 diffuse porous chronologies.

To test whether the negative effect of summer temperatures might offset an enhancement of growth by warmer spring temperatures, we tested for the joint effects of April and June-July $T_{max}$ on RWI.
Results were qualitatively similar to the univariate correlations (Fig. `r tree_rings`), with significant (at p = 0.05) positive correlations to April $T_{max}$ for only 4% of chronologies and significant negative correlations with June-July $T_{max}$ for 77% of chronologies (`r EDT_tree_ring_summary`).

@@

<!--summary, setting expectation of extra C-->
Together, our results demonstrate that warmer spring temperatures in the temperate deciduous forests of eastern North America advance the timing of stem growth but have little effect on total annual growth (Figs. `r schematic_summary`-`r tree_rings`). 
The observed advance in stem growth under warmer springs parallels advances observed for canopy foliage phenology (Fig. `r DOY_timing`a-b)[@keenan_net_2014;@finzi_carbon_2020;@stuble_plant_2021] and NEE[@keenan_net_2014;@finzi_carbon_2020]. 
However, inconsistent with the concept that an earlier start to growth would increase $ANPP_{woody}$, we demonstrate that warmer springs either hasten the deceleration of stem expansion or otherwise fail to translate extended growing seasons into biologically significant increases in stem growth (Fig. `r schematic_summary`), and thereby have negligible effect on total annual growth for most species and locations (Fig. `r tree_rings`).
Our observations suggest that the deceleration of stem expansion, which occurs mid-summer near the time of peak canopy greenness (`r EDF_growth_curves_all`)[@finzi_carbon_2020;@etzold_number_2022], is driven by cues other than photosynthate limitation, such as water stress, daylength, or sink limitation[@zweifel_why_2021;@etzold_number_2022;@zani_increased_2020;@keenan_timing_2015;<!--@fu_variation_2014;-->@delpierre_wood_2016;@delpierre_temperate_2016].
Combined with observations that warming springs tend to lengthen the season of C sequestration[@lu_no_2022;@finzi_carbon_2020] and increase net C sequestration[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014;@crabbe_extreme_2016;@zhou_explaining_2016;@fu_maximum_2019;@finzi_carbon_2020], our findings imply a lengthening of the period from peak stem growth to the cessation of CO~2~ uptake by the ecosystem and an increase in C allocated to functions other than stem expansion.
What happens with this extra C?

<!--extra C is probably not going to ANPP_woody-->
Additional C fixation in years with warm springs[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014;@crabbe_extreme_2016;@zhou_explaining_2016;@fu_maximum_2019;@finzi_carbon_2020] does not substantially augment stem diameter growth in the current or following year (`r EDF_tree_rings_TMAX_extended`), and is unlikely to augment $ANPP_{woody}$. 
Although $ANPP_{woody}$ is routinely calculated based on stem growth, the two can be partially decoupled through differences in wood density or C content[@delpierre_wood_2016].
It is theoretically possible that extra C is allocated to enhanced cell wall thickening, a process that lags behind stem expansion[@cuny_woody_2015], or to a higher ratio of high-density latewood to lower-density earlywood, resulting in formation of more C-dense wood in years with warm springs.
However, existing evidence indicates that warm springs have a neutral or negative effect on latewood width[<!--@tardif_influence_2006;-->@roibu_climatic_2020;@kern_multiple_2013], which is more strongly controlled by summer drought stress[<!--@tardif_influence_2006;-->@roibu_climatic_2020;@kern_multiple_2013], suggesting that a positive effect of warm springs on the total C content of annual rings is unlikely. 
Rather, the most likely explanation of the observed lack of correlation between spring temperatures and stem growth is that growth is sink-limited, such that increased GPP does not enhance $ANPP_{woody}$[@delpierre_wood_2016].

@@

<!--.short vs long-term effects-->
It is possible that as spring warming continues, forests will adjust to directional changes in growing season length with an enhancement of $ANPP_{woody}$.
Across latitudinal gradients, warmer spring temperatures are associated with earlier leaf-out[@delgado_differences_2020] and longer growing seasons, which in turn are correlated with greater tree growth[@anderson-teixeira_joint_2021], woody productivity[@banburymorgan_global_2021], and NEE[@churkina_spatial_2005]<!--Can we replace this churkina_2005 ref with another NEE ref? NO, it's making a different point.-->.<!-- The point here is that if you compare across sites in different climates (Harvard Forest and SCBI, for example), you do see greater stem growth associated with the longer growing seasons. Hopefully my rewording makes that clear. -->
Thus, warming spring temperatures are likely to increase the biophysical potential for annual tree growth.
If extra photosynthate made available through a growing difference between GPP and $ANPP_{woody}$ is allocated to functions that relieve limitations on woody growth -- for example, by enhancing nutrient and water acquisition through enhanced allocation to roots[@zohner_how_2021;@liu_phenological_2022] -- it is possible that warming spring temperatures could ultimately increase $ANPP_{woody}$ through indirect mechanisms.
Understanding how warming spring temperatures are influencing C allocation within ecosystems remains an important outstanding question.

<!--the dominant effect of temperature on ANPP_woody is a negative effect of gs temperature-->
Regardless of the influence of spring temperatures on C cycling within the ecosystem, our results clearly demonstrate that the dominant effects of temperature on deciduous tree growth occur not in the spring, but during the peak growing season, when increased atmospheric demand associated with high temperatures limits growth[@novick_increasing_2016;@delpierre_wood_2016;@etzold_number_2022;@zweifel_why_2021]. 
Indeed, the timing of peak growth in June and July (`r EDT_phenology_parameters`, `r EDF_growth_curves_all`) coincides with the timing of the greatest sensitivity of annual growth to $T_{max}$ (Fig. `r tree_rings`, `r EDT_tree_ring_summary`).
This finding is consistent with numerous tree-ring studies demonstrating strong sensitivity of growth to drought stress or high temperatures during the peak growing season[@dorangeville_drought_2018;@helcoski_growing_2019;@anderson-teixeira_joint_2021;@babst_twentieth_2019].
<!--*It is also worth noting that growth is often influenced by temperature of the preceding year(s) through its influence on production and storage of NSCs[@refs;@helcoski], and indeed we observe many significant effects of $T_{max}$ in the preceding growing season, albeit very few in the preceding spring (`r EDF_tree_rings_TMAX_extended`).*--><!--this sentence feels tangential, but discussion of these lag effects was requested by one reviewer.-->
Warm springs may also amplify summer drought stress in some times and places, effectively canceling out any positive effects of an extended growing period[@buermann_widespread_2018;@zhang_drought_2021;@lian_summer_2022].
While such an interaction was unlikely to have played a major role within the scope of our dendrometer band study, given relatively mesic conditions and lack of significant correlation between spring temperatures and summer drought stress (see Methods), our tree-ring analysis does reveal a higher frequency of negative than positive correlations of annual growth to spring temperatures, particularly for ring porous species in cooler climates (Fig. `r tree_rings`, `r EDT_tree_ring_summary`).
Thus, warm springs can have a net negative effect on growth at some times or places[@gao_earlier_2022].
